An introduction to wavelets using R

### 2022-02-06 text/14/wavelets.Rmd
Added 2-D images to wavelets.Rmd

```{r graphing-functions}
graph_with_axis <- function(x, lab, limits=range(x)) {
  par(mar=c(2.6, 0, 0, 0))
  plot((1:2048)/2048, x, 
       xlim=c(0, 1.2), 
       ylim=limits, 
       type="l",
       axes=FALSE)
  axis(side=1, at=(0:4)/4, labels=c(1, 512, 1024, 1536, 2048))
  text(1.1, x[2048], lab, adj=0.5)
}
graph_without_axis <- function(x, lab, limits=range(x)) {
  par(mar=c(0, 0, 0, 0))
  plot((1:2048)/2048, x, 
       xlim=c(0, 1.2), 
       ylim=c(limits[1]*1.1, limits[2]), # adjust lower limit
       type="l",
       axes=FALSE)
  text(1.1, x[2048], lab, adj=0.5)
}
needle_plot <- function(y, lab, x=seq(0, 1, along.with=y), limits=range(y)) {
  par(mar=c(0, 0, 0, 0))
  plot(
    x, y, 
    xlim=c(0, 1.2), 
    ylim=limits,
    type="n",
    axes=FALSE)
  segments(x, 0, x, y)
  text(1.1, 0, lab)
}
```

The head and tail function show the first few and the last few values for the signal. The graph looks like

@@

@@

The complex numbers in the Fourier decomposition are a short-hand for the mix of sines and cosines. The first number has only a real component and no imaginary component. The real component represent the mean of the entire signal.

@@

@@

```{r second-component}
ft.comp2 <- ft.ecg
ft.comp2[-2] <- 0
comp2 <- Re(fft(ft.comp2,inverse=TRUE))
graph_without_axis(sum1, "Using\nInverse fft")
{
  lines((1:2048)/2048, sum1, col="red", lty="dashed")
}
```

@@

```{r graph-third-component}
ft.comp3 <- ft.ecg
ft.comp3[-3] <- 0
comp3 <- Re(fft(ft.comp3,inverse=TRUE))
graph_without_axis(comp3, "Fourier\ncomponent")
```

@@

```{r graph-fourth-component}
ft.comp4 <- ft.ecg
ft.comp4[-4] <- 0
comp4 <- Re(fft(ft.comp4,inverse=TRUE))
graph_without_axis(comp4, "Fourier\ncomponent")
```

The fourth Fourier component has three peaks. The interesting thing is that the second and third and fourth Fourier components are orthogonal to one another. Actually, every Fourier component is orthogonal to every other Fourier component. Mathematical folks like "orthogonal" because it allows you to derive the mathematical properties of the decomposition with greater ease. For a practical person, what "orthogonal" means is that the information contained in one Fourier component does not "overlap" with that in any other Fourier component. It also means you can run the decomposition in any order (low frequency to high frequency, high frequency to low frequency, or even a haphazard frequency). Finally, orthogonality also preserves independence properties in certain settings.

@@

```{r thirty-second}
ft.comp32 <- ft.ecg
ft.comp32[-32] <- 0
comp32 <- Re(fft(ft.comp32,inverse=TRUE))
graph_without_axis(comp32, "Fourier\ncomponent")
```

@@

```{r count}
m <- max(comp32)
graph_without_axis(comp32, "Fourier\ncomponent", limits=c(-m, 1.2*m))
text((seq(1, 31, by=3)-0.5)/31, 1.2*m, seq(1, 31, by=3), col="red") 
```

@@

```{r rwt-package}
d4.rwt <- daubcqf(4)
d4.rwt
wt.rwt <- mdwt(ecg,d4.rwt$h.0,L=9)
names(wt.rwt)
head(wt.rwt$y)
tail(wt.rwt$y)
wt.rwt$L
iwt.rwt <- midwt(wt.rwt$y,d4.rwt$h.0,L=9)
names(iwt.rwt)
head(iwt.rwt$x)
tail(iwt.rwt$x)
iwt.rwt$L
```

@@

@@

```{r graph-d4}
wt.comp <- function(x, y) {
  single_wavelet <- as.numeric((1:2048)==y)
  as.vector(midwt(single_wavelet, d4.rwt$h.0, L=8)$x)
}
```

```{r calculate-functions}
calculate_mother <- function(i, L) {
  yl <- rep(0, 2048)
  yh <- matrix(0, nrow=2048, ncol=L)
  yh[i, L] <- 1
  as.vector(mirdwt(yl, yh, d4.rwt$h.0, L)$x)
}
calculate_father <- function(i, L) {
  yl <- rep(0, 2048)
  yh <- matrix(0, nrow=2048, ncol=L)
  yl[i] <- 1
  as.vector(mirdwt(yl, yh, d4.rwt$h.0, L)$x)
}
```

@@

```{r shifted}
{
  graph_without_axis(calculate_father(2^9-2^7+1, 7),"d4 father")
  graph_without_axis(calculate_father(2^9+1, 7),"d4 father")
  graph_without_axis(calculate_father(2^9+2^7+1, 7),"d4 father")
  graph_without_axis(calculate_mother(2^9-2^7+1, 7),"d4 mother")
  graph_without_axis(calculate_mother(2^9+1, 7),"d4 mother")
  graph_without_axis(calculate_mother(2^9+2^7+1, 7),"d4 mother")
}
```

@@

```{r wrapped}
{
  graph_without_axis(calculate_father(2^10+2^9+1+2^8+1, 7),"d4 father")
  graph_without_axis(calculate_mother(2^10+2^9+1+2^8+1, 7),"d4 mother")
}

@@

```{r stretch-and-squeeze}
{  
  graph_without_axis(calculate_father(2^9+1, 8),"d4 father")
  graph_without_axis(calculate_father(2^9+1, 7),"d4 father")
  graph_without_axis(calculate_father(2^9+1, 6),"d4 father")
  graph_without_axis(calculate_mother(2^9+1, 8),"d4 mother")
  graph_without_axis(calculate_mother(2^9+1, 7),"d4 mother")
  graph_without_axis(calculate_mother(2^9+1, 6),"d4 mother")
}
```

@@

```{r most-extreme}
{
  graph_without_axis(calculate_father(2^9+1, 1),"d4 father")
  graph_without_axis(calculate_mother(2^9+1, 1),"d4 mother")
}
```

@@

```{r high-pass}
g <- d4.rwt$h.0
g
```

@@

```{r low-pass}
h <- -d4.rwt$h.1
h
```

@@

```{r first-level-wavelet}
wt.1 <- mdwt(ecg, g, L=1)
V1 <- wt.1$y[   1:1024]
W1 <- wt.1$y[1025:2048]
{
  needle_plot(V1, "V1", limits=range(ecg))
  needle_plot(W1, "W1", limits=range(ecg))
  graph_without_axis(ecg, "Original\nsignal", limits=range(ecg))
}
```

@@

```{r second-level-wavelet}
wt.2 <- mdwt(ecg, g, L=2)
V2 <- wt.2$y[1:512]
W2 <- wt.2$y[513:1024]
quality_check_W1 <- wt.2$y[1025:2048]
all(quality_check_W1==W1)
{
  needle_plot(V2, "V2", limits=range(ecg))
  needle_plot(W2, "W2", limits=range(ecg))
  needle_plot(W1, "W1", limits=range(ecg))
  graph_without_axis(ecg, "Original\nsignal", limits=range(ecg))
}
```

@@

```{r transform-V1}
quality_check <- mdwt(V1, g, L=1)
quality_check_V2 <- quality_check$y[  1: 512]
quality_check_W2 <- quality_check$y[513:1024]
all(quality_check_V2==V2)
all(quality_check_W2==W2)
```

@@

```{r third-level-wavelet}
wt.3 <- mdwt(ecg, g, L=3)
V3 <- wt.3$y[1:256]
W3 <- wt.3$y[257:512]
W2_check3 <- wt.3$y[513:1024]
W1_check3 <- wt.3$y[1025:2048]
all(W2_check3==W2)
all(W1_check3==W1)
{
  needle_plot(V3, "V3", limits=range(ecg))
  needle_plot(W3, "W3", limits=range(ecg))
  needle_plot(W2, "W2", limits=range(ecg))
  needle_plot(W1, "W1", limits=range(ecg))
  graph_without_axis(ecg, "Original\nsignal", limits=range(ecg))
}
```

@@

```{r ninth-level-wavelet}
wt.9 <- mdwt(ecg, g, L=9)
V9 <- wt.9$y[   1:   4]
W9 <- wt.9$y[   5:   8]
W8 <- wt.9$y[   9:  16]
W7 <- wt.9$y[  17:  32]
W6 <- wt.9$y[  33:  64]
W5 <- wt.9$y[  65: 128]
W4 <- wt.9$y[ 129: 256]
W3 <- wt.9$y[ 257: 512]
W2 <- wt.9$y[ 513:1024]
W1 <- wt.9$y[1025:2048]
{
  needle_plot(V9, "V9", limits=range(ecg))
  needle_plot(W9, "W9", limits=range(ecg))
  needle_plot(W8, "W8", limits=range(ecg))
  needle_plot(W7, "W7", limits=range(ecg))
  needle_plot(W6, "W6", limits=range(ecg))
  needle_plot(W5, "W5", limits=range(ecg))
  needle_plot(W4, "W4", limits=range(ecg))
  needle_plot(W3, "W3", limits=range(ecg))
  needle_plot(W2, "W2", limits=range(ecg))
  needle_plot(W1, "W1", limits=range(ecg))
  graph_without_axis(ecg, "Original\nsignal", limits=range(ecg))
}
```

@@

First you dilute the W and V values at the top of the pyramid. Then you compute a running dot product of the reversed g vector with the diluted V9 values plus a running dot product of the reversed h vector with the diluted W9 values.

@@

wt.8 <- mdwt(ecg, g, L=8)
V8 <- wt.8$y[1:8]

@@

wt.7 <- mdwt(ecg, g, L=7)
V7 <- wt.7$y[1:16]

@@

wt.6 <- mdwt(ecg, g, L=6)
V6 <- wt.6$y[1:32]

@@

V9_coefficients <- rep(0, 2048)
V9_coefficients[1:4] <- V9
V9_signal <- as.vector(midwt(V9_coefficients, d4.rwt$h.0, L=9)$x)

V8_coefficients <- rep(0, 2048)
V8_coefficients[1:8] <- V8
V8_signal <- as.vector(midwt(V8_coefficients, d4.rwt$h.0, L=8)$x)

V7_coefficients <- rep(0, 2048)
V7_coefficients[1:16] <- V7
V7_signal <- as.vector(midwt(V7_coefficients, d4.rwt$h.0, L=7)$x)

V6_coefficients <- rep(0, 2048)
V6_coefficients[1:32] <- V6
V6_signal <- as.vector(midwt(V6_coefficients, d4.rwt$h.0, L=6)$x)

wt.5 <- mdwt(ecg, g, L=5)
V5 <- wt.5$y[1:64] 
V5_coefficients <- rep(0, 2048)
V5_coefficients[1:64] <- V5
V5_signal <- as.vector(midwt(V5_coefficients, d4.rwt$h.0, L=5)$x)

wt.4 <- mdwt(ecg, g, L=4)
V4 <- wt.4$y[1:128] 
V4_coefficients <- rep(0, 2048)
V4_coefficients[1:128] <- V4
V4_signal <- as.vector(midwt(V4_coefficients, d4.rwt$h.0, L=4)$x)

wt.3 <- mdwt(ecg, g, L=3)
V3 <- wt.3$y[1:256] 
V3_coefficients <- rep(0, 2048)
V3_coefficients[1:256] <- V3
V3_signal <- as.vector(midwt(V3_coefficients, d4.rwt$h.0, L=3)$x)

wt.2 <- mdwt(ecg, g, L=2)
V2 <- wt.2$y[1:512] 
V2_coefficients <- rep(0, 2048)
V2_coefficients[1:512] <- V2
V2_signal <- as.vector(midwt(V2_coefficients, d4.rwt$h.0, L=2)$x)

wt.1 <- mdwt(ecg, g, L=1)
V1 <- wt.1$y[1:1024] 
V1_coefficients <- rep(0, 2048)
V1_coefficients[1:1024] <- V1
V1_signal <- as.vector(midwt(V1_coefficients, d4.rwt$h.0, L=1)$x)

@@

W9_coefficients <- rep(0, 2048)
W9_coefficients[5:8] <- W9
W9_signal <- as.vector(midwt(W9_coefficients, d4.rwt$h.0, L=9)$x)

W8_coefficients <- rep(0, 2048)
W8_coefficients[9:16] <- W8
W8_signal <- as.vector(midwt(W8_coefficients, d4.rwt$h.0, L=9)$x)

W7_coefficients <- rep(0, 2048)
W7_coefficients[17:32] <- W7
W7_signal <- as.vector(midwt(W7_coefficients, d4.rwt$h.0, L=9)$x)

W6_coefficients <- rep(0, 2048)
W6_coefficients[33:64] <- W6
W6_signal <- as.vector(midwt(W6_coefficients, d4.rwt$h.0, L=9)$x)

W5_coefficients <- rep(0, 2048)
W5_coefficients[65:128] <- W5
W5_signal <- as.vector(midwt(W5_coefficients, d4.rwt$h.0, L=9)$x)

W4_coefficients <- rep(0, 2048)
W4_coefficients[129:256] <- W4
W4_signal <- as.vector(midwt(W4_coefficients, d4.rwt$h.0, L=9)$x)

W3_coefficients <- rep(0, 2048)
W3_coefficients[257:512] <- W3
W3_signal <- as.vector(midwt(W3_coefficients, d4.rwt$h.0, L=9)$x)

W2_coefficients <- rep(0, 2048)
W2_coefficients[513:1024] <- W2
W2_signal <- as.vector(midwt(W2_coefficients, d4.rwt$h.0, L=9)$x)

W1_coefficients <- rep(0, 2048)
W1_coefficients[1025:2048] <- W1
W1_signal <- as.vector(midwt(W1_coefficients, d4.rwt$h.0, L=9)$x)

@@

@@

img <- read.table("klaus256.txt")
img <- 2*img-1

img.plot <- function(imat,n=256) {
xsq <- c(0,0,1,1)
ysq <- c(0,1,1,0)
par(mar=rep(0,4),xaxs="i",yaxs="i")
plot(c(1,n+1),c(1,n+1),type="n",axes=FALSE)
for (i in 1:n) {
for (j in 1:n){
d <- 0.5*imat[n+1-j,i]+0.5
d <- min(1,max(0,d))
polygon(xsq+i,ysq+j,density=-1,border=NA,col=rgb(d,d,d))
}
}
}
img.plot(img)

Dog image

@@

g <- daubcqf(8)$h.0
g
[1] 0.23037781 0.71484657 0.63088077 -0.02798377 -0.18703481 0.03084138
[7] 0.03288301 -0.01059740
h <- -daubcqf(8)$h.1
h
[1] -0.01059740 -0.03288301 0.03084138 0.18703481 -0.02798377 -0.63088077
[7] 0.71484657 -0.23037781

@@

ecg
Axis

@@

wt.img.1 <- mdwt(img, g, L=1)
sum(img[1:8,1:8]*outer(g,g)),wt.img.1$y[ 1, 1],"V1")
V1
By hand -1.580973
Using R -1.580973
sum(img[1:8,1:8]*outer(h,h)),wt.img.1$y[129,129],"W1")
W1
By hand -0.002825861
Using R -0.002825861
sum(img[1:8,1:8]*outer(h,g)),wt.img.1$y[129, 1],"W1")
W1
By hand -0.01225749
Using R -0.01225749
sum(img[1:8,1:8]*outer(g,h)),wt.img.1$y[ 1,129],"W1")
W1
By hand -0.005042734
Using R -0.005042734

@@

Wavelet coefficients arrangement

@@

>>>>

```{r graphing-functions}
graph_with_axis <- function(x, lab, limits=range(x)) {
  par(mar=c(2.6, 0, 0, 0))
  plot((1:2048)/2048, x, 
       xlim=c(0, 1.2), 
       ylim=limits, 
       type="l",
       axes=FALSE)
  axis(side=1, at=(0:4)/4, labels=c(1, 512, 1024, 1536, 2048))
  text(1.1, x[2048], lab, adj=0.5, cex=0.67)
}
graph_without_axis <- function(x, lab, limits=range(x)) {
  par(mar=c(0, 0, 0, 0))
  plot((1:2048)/2048, x, 
       xlim=c(0, 1.2), 
       ylim=c(limits[1], limits[2]),
       type="l",
       axes=FALSE)
  text(1.1, x[2048], lab, adj=0.5, cex=0.67)
}
needle_plot <- function(y, lab, x=seq(0, 1, along.with=y), limits=range(y)) {
  par(mar=c(0, 0, 0, 0))
  plot(
    x, y, 
    xlim=c(0, 1.2), 
    ylim=limits,
    type="n",
    axes=FALSE)
  segments(x, 0, x, y)
  text(1.1, 0, lab, cex=0.67)
}
```

### The ecg signal

Here's a graph of data representing an electrocardiogram (ecg) signal that will help illustrate the use of Fourier transforms and wavelets.

@@

### Components of the Fourier transform

@@

The complex numbers in the Fourier decomposition are a short-hand for the mix of sines and cosines. The first number has only a real component and no imaginary component. This component represent the mean of the entire signal.

@@

### Zeroing out components

@@

```{r second-component}
ft.comp2 <- ft.ecg
ft.comp2[-2] <- 0
comp2 <- Re(fft(ft.comp2,inverse=TRUE))
graph_without_axis(sum1, "Using\nInverse fft")
{
  lines((1:2048)/2048, sum1, col="red", lty="dashed", lwd=3)
}
```

@@

```{r graph-third-component}
ft.comp3 <- ft.ecg
ft.comp3[-3] <- 0
comp3 <- Re(fft(ft.comp3,inverse=TRUE))
graph_without_axis(comp3, "Third\ncomponent")
```

@@

```{r graph-fourth-component}
ft.comp4 <- ft.ecg
ft.comp4[-4] <- 0
comp4 <- Re(fft(ft.comp4,inverse=TRUE))
graph_without_axis(comp4, "Fourth\ncomponent")
```

The fourth Fourier component has three peaks.

### Orthogonality

The interesting thing is that the second and third and fourth Fourier components are orthogonal to one another. Actually, every Fourier component is orthogonal to every other Fourier component. Mathematical folks like "orthogonal" because it allows you to derive the mathematical properties of the decomposition with greater ease. For a practical person, what "orthogonal" means is that the information contained in one Fourier component does not "overlap" with that in any other Fourier component. It also means you can run the decomposition in any order (low frequency to high frequency, high frequency to low frequency, or even a haphazard frequency). Finally, orthogonality also preserves independence properties in certain settings.

@@

```{r thirty-second}
ft.comp32 <- ft.ecg
ft.comp32[-32] <- 0
comp32 <- Re(fft(ft.comp32,inverse=TRUE))
graph_without_axis(comp32, "32nd\ncomponent")
```

@@

```{r count}
m <- max(comp32)
graph_without_axis(comp32, "32nd\ncomponent", limits=c(-m, 1.2*m))
text((seq(1, 31, by=3)-0.5)/31, 1.2*m, seq(1, 31, by=3), col="red", cex=0.67) 
```

@@

```{r rwt-package}
d4.rwt <- daubcqf(4)
d4.rwt
wt.rwt <- mdwt(ecg,daubcqf(4)$h.0,L=9)
names(wt.rwt)
head(wt.rwt$y)
tail(wt.rwt$y)
wt.rwt$L
iwt.rwt <- midwt(wt.rwt$y,daubcqf(4)$h.0,L=9)
names(iwt.rwt)
head(iwt.rwt$x)
tail(iwt.rwt$x)
iwt.rwt$L
```

@@

### wavelets package

@@

```{r graph-d4}
wt.comp <- function(x, y) {
  single_wavelet <- as.numeric((1:2048)==y)
  as.vector(midwt(single_wavelet, daubcqf(4)$h.0, L=8)$x)
}
```

```{r calculate-functions}
calculate_mother <- function(i, L) {
  yl <- rep(0, 2048)
  yh <- matrix(0, nrow=2048, ncol=L)
  yh[i, L] <- 1
  as.vector(mirdwt(yl, yh, daubcqf(4)$h.0, L)$x)
}
calculate_father <- function(i, L) {
  yl <- rep(0, 2048)
  yh <- matrix(0, nrow=2048, ncol=L)
  yl[i] <- 1
  as.vector(mirdwt(yl, yh, daubcqf(4)$h.0, L)$x)
}
```

@@

```{r shifted}
{
  graph_without_axis(calculate_father(2^9-2^7+1, 7),"d4 father\nshifted left")
  graph_without_axis(calculate_father(2^9+2^7+1, 7),"d4 father\nshifted right")
  graph_without_axis(calculate_mother(2^9-2^7+1, 7),"d4 mother\nshifted left")
  graph_without_axis(calculate_mother(2^9+2^7+1, 7),"d4 mother\nshifted right")
}
```

@@

```{r wrapped}
{
  graph_without_axis(calculate_father(2^10+2^9+1+2^8+1, 7),"d4 father, wrapped")
  graph_without_axis(calculate_mother(2^10+2^9+1+2^8+1, 7),"d4 mother, wrapped")
}

@@

```{r stretch-and-squeeze}
{  
  graph_without_axis(calculate_father(2^9+1, 8),"d4 father\nstretched")
  graph_without_axis(calculate_father(2^9+1, 6),"d4 father\nsqueezed")
  graph_without_axis(calculate_mother(2^9+1, 8),"d4 mother\nstretched")
  graph_without_axis(calculate_mother(2^9+1, 6),"d4 mother\nsqueezed")
}
```

@@

```{r most-extreme}
{
  graph_without_axis(calculate_father(2^9+1, 1),"d4 father\nmost extreme")
  graph_without_axis(calculate_mother(2^9+1, 1),"d4 mother\nmost extreme")
}
```

@@

```{r high-pass}
g <- daubcqf(4)$h.0
g
```

@@

```{r low-pass}
h <- -daubcqf(4)$h.1
h
```

@@

```{r first-level-wavelet}
wt.1 <- mdwt(ecg, daubcqf(4)$h.0, L=1)
V1 <- wt.1$y[   1:1024]
W1 <- wt.1$y[1025:2048]
{
  needle_plot(V1, "V1", limits=range(ecg))
  needle_plot(W1, "W1", limits=range(ecg))
  graph_without_axis(ecg, "Original\nsignal", limits=range(ecg))
}
```

@@

```{r second-level-wavelet}
wt.2 <- mdwt(ecg, daubcqf(4)$h.0, L=2)
V2 <- wt.2$y[1:512]
W2 <- wt.2$y[513:1024]
quality_check_W1 <- wt.2$y[1025:2048]
all(quality_check_W1==W1)
{
  needle_plot(V2, "V2", limits=range(ecg))
  needle_plot(W2, "W2", limits=range(ecg))
  needle_plot(W1, "W1", limits=range(ecg))
  graph_without_axis(ecg, "Original\nsignal", limits=range(ecg))
}
```

@@

```{r transform-V1}
quality_check <- mdwt(V1, daubcqf(4)$h.0, L=1)
quality_check_V2 <- quality_check$y[  1: 512]
quality_check_W2 <- quality_check$y[513:1024]
all(quality_check_V2==V2)
all(quality_check_W2==W2)
```

@@

```{r third-level-wavelet}
wt.3 <- mdwt(ecg, daubcqf(4)$h.0, L=3)
V3 <- wt.3$y[1:256]
W3 <- wt.3$y[257:512]
W2_check3 <- wt.3$y[513:1024]
W1_check3 <- wt.3$y[1025:2048]
all(W2_check3==W2)
all(W1_check3==W1)
{
  needle_plot(V3, "V3", limits=range(ecg))
  needle_plot(W3, "W3", limits=range(ecg))
  needle_plot(W2, "W2", limits=range(ecg))
  needle_plot(W1, "W1", limits=range(ecg))
  graph_without_axis(ecg, "Original\nsignal", limits=range(ecg))
}
```

@@

```{r ninth-level-wavelet}
wt.9 <- mdwt(ecg, daubcqf(4)$h.0, L=9)
V9 <- wt.9$y[   1:   4]
W9 <- wt.9$y[   5:   8]
W8 <- wt.9$y[   9:  16]
W7 <- wt.9$y[  17:  32]
W6 <- wt.9$y[  33:  64]
W5 <- wt.9$y[  65: 128]
W4 <- wt.9$y[ 129: 256]
W3 <- wt.9$y[ 257: 512]
W2 <- wt.9$y[ 513:1024]
W1 <- wt.9$y[1025:2048]
{
  needle_plot(V9, "V9", limits=range(ecg))
  needle_plot(W9, "W9", limits=range(ecg))
  needle_plot(W8, "W8", limits=range(ecg))
  needle_plot(W7, "W7", limits=range(ecg))
  needle_plot(W6, "W6", limits=range(ecg))
  needle_plot(W5, "W5", limits=range(ecg))
  needle_plot(W4, "W4", limits=range(ecg))
  needle_plot(W3, "W3", limits=range(ecg))
  needle_plot(W2, "W2", limits=range(ecg))
  needle_plot(W1, "W1", limits=range(ecg))
  graph_without_axis(ecg, "Original\nsignal", limits=range(ecg))
}
```

@@

First you dilute the W and V values at the top of the pyramid. Then you compute a running dot product of the reversed g vector with the diluted V9 values plus a running dot product of the reversed h vector with the diluted W9 values. Please note that using the reversed order for the g and h vectors is unique to the Daubechies family of wavelets. Other wavelets will use an entirely different set of vectors in the rolling dot product.

![Figure 3. Matrix representation of reconstruction](http://www.pmean.com/new-images/14/wavelets-03.png)

@@

wt.8 <- mdwt(ecg, daubcqf(4)$h.0, L=8)
V8 <- wt.8$y[1:8]

@@

wt.7 <- mdwt(ecg, daubcqf(4)$h.0, L=7)
V7 <- wt.7$y[1:16]

@@

wt.6 <- mdwt(ecg, daubcqf(4)$h.0, L=6)
V6 <- wt.6$y[1:32]

@@

V9_coefficients <- rep(0, 2048)
V9_coefficients[1:4] <- V9
V9_signal <- as.vector(midwt(V9_coefficients, daubcqf(4)$h.0, L=9)$x)

V8_coefficients <- rep(0, 2048)
V8_coefficients[1:8] <- V8
V8_signal <- as.vector(midwt(V8_coefficients, daubcqf(4)$h.0, L=8)$x)

V7_coefficients <- rep(0, 2048)
V7_coefficients[1:16] <- V7
V7_signal <- as.vector(midwt(V7_coefficients, daubcqf(4)$h.0, L=7)$x)

V6_coefficients <- rep(0, 2048)
V6_coefficients[1:32] <- V6
V6_signal <- as.vector(midwt(V6_coefficients, daubcqf(4)$h.0, L=6)$x)

wt.5 <- mdwt(ecg, daubcqf(4)$h.0, L=5)
V5 <- wt.5$y[1:64] 
V5_coefficients <- rep(0, 2048)
V5_coefficients[1:64] <- V5
V5_signal <- as.vector(midwt(V5_coefficients, daubcqf(4)$h.0, L=5)$x)

wt.4 <- mdwt(ecg, daubcqf(4)$h.0, L=4)
V4 <- wt.4$y[1:128] 
V4_coefficients <- rep(0, 2048)
V4_coefficients[1:128] <- V4
V4_signal <- as.vector(midwt(V4_coefficients, daubcqf(4)$h.0, L=4)$x)

wt.3 <- mdwt(ecg, daubcqf(4)$h.0, L=3)
V3 <- wt.3$y[1:256] 
V3_coefficients <- rep(0, 2048)
V3_coefficients[1:256] <- V3
V3_signal <- as.vector(midwt(V3_coefficients, daubcqf(4)$h.0, L=3)$x)

wt.2 <- mdwt(ecg, daubcqf(4)$h.0, L=2)
V2 <- wt.2$y[1:512] 
V2_coefficients <- rep(0, 2048)
V2_coefficients[1:512] <- V2
V2_signal <- as.vector(midwt(V2_coefficients, daubcqf(4)$h.0, L=2)$x)

wt.1 <- mdwt(ecg, daubcqf(4)$h.0, L=1)
V1 <- wt.1$y[1:1024] 
V1_coefficients <- rep(0, 2048)
V1_coefficients[1:1024] <- V1
V1_signal <- as.vector(midwt(V1_coefficients, daubcqf(4)$h.0, L=1)$x)

@@

W9_coefficients <- rep(0, 2048)
W9_coefficients[5:8] <- W9
W9_signal <- as.vector(midwt(W9_coefficients, daubcqf(4)$h.0, L=9)$x)

W8_coefficients <- rep(0, 2048)
W8_coefficients[9:16] <- W8
W8_signal <- as.vector(midwt(W8_coefficients, daubcqf(4)$h.0, L=9)$x)

W7_coefficients <- rep(0, 2048)
W7_coefficients[17:32] <- W7
W7_signal <- as.vector(midwt(W7_coefficients, daubcqf(4)$h.0, L=9)$x)

W6_coefficients <- rep(0, 2048)
W6_coefficients[33:64] <- W6
W6_signal <- as.vector(midwt(W6_coefficients, daubcqf(4)$h.0, L=9)$x)

W5_coefficients <- rep(0, 2048)
W5_coefficients[65:128] <- W5
W5_signal <- as.vector(midwt(W5_coefficients, daubcqf(4)$h.0, L=9)$x)

W4_coefficients <- rep(0, 2048)
W4_coefficients[129:256] <- W4
W4_signal <- as.vector(midwt(W4_coefficients, daubcqf(4)$h.0, L=9)$x)

W3_coefficients <- rep(0, 2048)
W3_coefficients[257:512] <- W3
W3_signal <- as.vector(midwt(W3_coefficients, daubcqf(4)$h.0, L=9)$x)

W2_coefficients <- rep(0, 2048)
W2_coefficients[513:1024] <- W2
W2_signal <- as.vector(midwt(W2_coefficients, daubcqf(4)$h.0, L=9)$x)

W1_coefficients <- rep(0, 2048)
W1_coefficients[1025:2048] <- W1
W1_signal <- as.vector(midwt(W1_coefficients, daubcqf(4)$h.0, L=9)$x)

@@

### Applications of wavelets

### Two dimensional wavelets

@@

```{r read-dog-image}
knitr::opts_chunk$set(fig.width=5, fig.height=5)
fn <- "https://www2.isye.gatech.edu/isyebayes/data/klaus256.asc"
img <- read.table(fn)
img <- 2*img-1
```

```{r display-dog-image}
img.plot <- function(imat,n=256) {
  xsq <- c(0,0,1,1)
  ysq <- c(0,1,1,0)
  par(mar=rep(0,4),xaxs="i",yaxs="i")
  plot(c(1,n+1),c(1,n+1),type="n",axes=FALSE)
  # imat <- imat/max(abs(imat))
  for (i in 1:n) {
    for (j in 1:n){
      d <- 0.5*imat[n+1-j,i]+0.5
      d <- min(1,max(0,d))
      polygon(xsq+i,ysq+j,density=-1,border=NA,col=rgb(d,d,d))
    }
  }
}

img.plot(img)
```

@@

```{r d8-wavelet-g}
g <- daubcqf(8)$h.0
g
```

```{r d8-wavelet-h}
h <- -daubcqf(8)$h.1
h
```

@@

```{r d8-in-1d, fig.height=1}
d8_father_coefficients <- rep(0, 2048)
d8_father_coefficients[3] <- 1
d8_father <- as.vector(midwt(d8_father_coefficients, daubcqf(8)$h.0, L=8)$x)
d8_mother_coefficients <- rep(0, 2048)
d8_mother_coefficients[10] <- 1
d8_mother <- as.vector(midwt(d8_mother_coefficients, daubcqf(8)$h.0, L=8)$x)
{
  graph_without_axis(d8_father, "D8 father")
  graph_without_axis(d8_mother, "D8 mother")
}
```

@@

```{r decompose-and-restructure}
wt.img.4 <- mdwt(img, daubcqf(8)$h.0, L=4)
reconstructed.img <- midwt(wt.img.4$y, daubcqf(8)$h.0, L=4)
img.plot(reconstructed.img$x)
```

```{r check-2d-calculations}
wt.img.1 <- mdwt(img, daubcqf(8)$h.0, L=1)
c(sum(img[1:8,1:8]*outer(g,g)), wt.img.1$y[ 1,   1])
c(sum(img[1:8,1:8]*outer(g,h)),wt.img.1$y[  1, 129])
c(sum(img[1:8,1:8]*outer(h,g)),wt.img.1$y[129,   1])
c(sum(img[1:8,1:8]*outer(h,h)),wt.img.1$y[129, 129])
```

@@

### Wavelet coefficients arrangement

@@

```{r level-1-image-g-g}
gg1 <- matrix(0, nrow=256, ncol=256)
gg1[1:128, 1:128] <- wt.img.1$y[1:128, 1:128]
gg1_signal <- midwt(gg1, daubcqf(8)$h.0, L=1)  
img.plot(gg1_signal$x)
```

```{r level-1-image-g-h}
gh1 <- matrix(0, nrow=256, ncol=256)
gh1[1:128, 129:256] <- wt.img.1$y[1:128, 129:256]
gh1_signal <- midwt(gh1, daubcqf(8)$h.0, L=1)  
img.plot(gh1_signal$x)
```

```{r level-1-image-h-g}
hg1 <- matrix(0, nrow=256, ncol=256)
hg1[1:128, 129:256] <- wt.img.1$y[129:256, 1:128]
hg1_signal <- midwt(hg1, daubcqf(8)$h.0, L=1)  
img.plot(hg1_signal$x)
```

```{r level-1-image-h-h}
hh1 <- matrix(0, nrow=256, ncol=256)
hh1[129:256, 129:256] <- wt.img.1$y[129:256, 129:256]
hh1_signal <- midwt(hh1, daubcqf(8)$h.1, L=1)  
img.plot(hh1_signal$x)
```

```{r level-4-image-g-g}
gg4 <- matrix(0, nrow=256, ncol=256)
gg4[1:16, 1:16] <- wt.img.4$y[1:16, 1:16]
gg4_signal <- midwt(gg4, daubcqf(8)$h.0, L=4)  
img.plot(gg4_signal$x)
```

```{r level-4-image-g-h}
gh4 <- matrix(0, nrow=256, ncol=256)
gh4[1:16, 17:32] <- wt.img.4$y[1:16, 17:32]
gh4[1:32, 33:64] <- wt.img.4$y[1:32, 33:64]
gh4[1:64, 65:128] <- wt.img.4$y[1:64, 65:128]
gh4[1:128,129:256] <- wt.img.4$y[1:128, 129:256]
gh4_signal <- midwt(gh4, daubcqf(8)$h.0, L=4)  
img.plot(gh4_signal$x)
```

```{r level-4-image-h-g}
hg4 <- matrix(0, nrow=256, ncol=256)
hg4[17:32, 1:16] <- wt.img.4$y[17:32, 1:16]
hg4[33:64, 1:32] <- wt.img.4$y[33:64, 1:32]
hg4[65:128, 1:64] <- wt.img.4$y[65:128, 1:64]
hg4[129:256, 1:128] <- wt.img.4$y[129:256, 1:128]
hg4_signal <- midwt(hg4, daubcqf(8)$h.0, L=4)  
img.plot(hg4_signal$x)
```

```{r level-4-image-h-h}
hh4 <- matrix(0, nrow=256, ncol=256)
hh4[17:32, 17:32] <- wt.img.4$y[17:32, 17:32]
hh4[33:64, 33:64] <- wt.img.4$y[33:64, 33:64]
hh4[65:128, 65:128] <- wt.img.4$y[65:128, 65:128]
hh4[129:256, 129:256] <- wt.img.4$y[129:256, 129:256]
hh4_signal <- midwt(hh4, daubcqf(8)$h.0, L=4)  
img.plot(hh4_signal$x)
```

```{r level-4-detail}
detail4 <- gh4 + hg4 + hh4
detail4_signal <- midwt(detail4, daubcqf(8)$h.0, L=4)  
img.plot(detail4_signal$x)
```

```{r compress}
pctl_99 <- quantile(abs(wt.img.4$y), probs=0.99)
top_01 <- matrix(0, nrow=256, ncol=256)
top_01[abs(wt.img.4$y)>pctl_99] <- wt.img.4$y[abs(wt.img.4$y)>pctl_99]
top_01_image <- midwt(top_01, daubcqf(8)$h.0, L=4)
img.plot(top_01_image$x)
